### 2024-05-31 journal/response.txt

We added the reminder.  Since q stands for queue, we thought q would remind readers of the queue's size.

> - Sec 3.4, and later. At many places operations are written in different font styles. Please follow a uniform format.

Done.

> - ln 845-848, right. I got confused here. Lemma 4 is only about the indexes, isn't it?

Since expression (3.3) defines subblocks in terms of indexes, we can apply Lemma 4.  Some more explanation has been added to the proof.

> - ln 1084-1088, right. Here do you mean that L is a valid sequential execution of the queue.
> 
> - Lemma 15. After reading the proof I understood that legal linearization ordering only  means respecting real-time order. This is a concrete example why I think a preliminaries section might help.
> 
> - ln 1192-1993, right. Please formalize the notion of time in an execution,
> 
> - Theorem 22. What exactly is the size of the queue?
> 
> - Sec 6. Why do you need read-back trees? How is that they facilitate the task of keeping memory bounded? Why the first algorithm cannot be modified to remove bocks, let us say using some sort of epoch-based memory reclamation scheme?

As mentioned above, we could not figure out a simpler way to do this.

> - ln 1279-1283, left. Why? A whole path might need to be copied, isn’t it?

Since all accesses are through the root, one cannot reach a node without traversing all nodes along the path from the root to that node.  Copying those nodes then only increases the running time by a constant factor.  (The sentence has been reworded to avoid confusion.)

> - ln 1442. Why not Help(), in accordance to Dequeue(), for example?

We removed the parentheses from Dequeue for consistency.

> - Sec 7. Your algorithms are adaptive on the size of the queue but not on the number of processes. I wonder if an interesting future direction would be to make your algorithms fully adaptive.

Yes, this was mentioned in sentence 3 of Sec 7.

### 2024-05-31 journal/boundedspace.tex
working on reviewer 2's comments

Known lock-free search trees have step complexity that includes a term linear in $p$ \cite{EFHR14,Ko20}.  
However, we do not require all the standard search tree operations.
Instead of a standard insertion, we allow a \op{Refresh}'s insertion to fail if another
concurrent \op{Refresh} succeeds in inserting a block, just as the \op{CAS} on line \ref{cas}
can fail if a concurrent \op{Refresh} does a successful \op{CAS}.
Moreover, the insertion should succeed only if the newly inserted block has a larger index than any other block in the RBT.
Thus, we can use a particularly simple concurrent RBT implementation.
A sequential RBT can be made persistent using the classic node-copying technique of 
Driscoll et al.~\cite{DSST89}:  all RBT nodes are immutable, and operations on the 
RBT make a new copy of each RBT node $x$ that must be modified, as well
as each RBT node along the path from the RBT's root to~$x$.
The RBT reachable from the new copy of the root is the result of applying the RBT operation.
This  adds only a constant factor to the running time of any routine designed for a (sequential) RBT.
Once a process has performed an update to the RBT representing the blocks of a node 
$\var{v}$ in the \ordering\ tree, 
it uses a CAS to swing $\var{v}$'s pointer from the previous RBT root to the new RBT root.
A search in the RBT can simply read the pointer to the RBT root and perform a standard
sequential search on it.
Bashari and Woelfel ~\cite{DBLP:conf/podc/BashariW21} used persistent RBTs in a similar way for a snapshot data structure.

>>>>

Known lock-free search trees have step complexity that includes a term linear in $p$ \cite{EFHR14,Ko20}.  
However, we do not require all the standard search tree operations.
Instead of a standard insertion, we allow a \op{Refresh}'s insertion to fail if another
concurrent \op{Refresh} succeeds in inserting a block, just as the \op{CAS} on line \ref{cas}
can fail if a concurrent \op{Refresh} does a successful \op{CAS}.
Moreover, the insertion should succeed only if the newly inserted block has a larger index than any other block in the RBT.
Thus, we can use a particularly simple concurrent RBT implementation.
A sequential RBT can be made persistent using the classic node-copying technique of 
Driscoll et al.~\cite{DSST89}:  all RBT nodes are immutable, and operations on the 
RBT make a new copy of each RBT node $x$ that must be modified, as well
as each RBT node along the path from the RBT's root to~$x$.
The RBT reachable from the new copy of the root is the result of applying the RBT operation.
This \rev{increases the running time of any routine designed for a (sequential) RBT only by a constant factor, assuming that nodes are accessed only by following pointers from the root (as is the case for all RBT operations we require)}.
Once a process has performed an update to the RBT representing the blocks of a node 
$\var{v}$ in the \ordering\ tree, 
it uses a CAS to swing $\var{v}$'s pointer from the previous RBT root to the new RBT root.
A search in the RBT can simply read the pointer to the RBT root and perform a standard
sequential search on it.
Bashari and Woelfel ~\cite{DBLP:conf/podc/BashariW21} used persistent RBTs in a similar way for a snapshot data structure.

### 2024-05-31 journal/correctness.tex
working on reviewer 2's comments

\begin{lemma} \label{lem::subblocksDistinct}
If $B$ and $B'$ are two \rev{different} blocks in nodes at the same depth in the ordering tree, their sets of subblocks are disjoint.
\end{lemma}
\begin{proof}
We prove the lemma by reverse induction on the depth.
If $B$ and $B'$ are in leaves, they have no subblocks, so the claim holds.
Assume the claim holds for nodes at depth $d+1$ and let $B$ and $B'$ be two blocks in nodes at depth $d$.
Consider the direct subblocks of $B$ and $B'$ defined by~(\ref{defsubblock}).
If $B$ and $B'$ are in different nodes at depth $d$, then their direct subblocks are disjoint.
If $B$ and $B'$ are in the same node, it follows from Lemma \ref{lem::headProgress} that their direct subblocks are disjoint.
Either way, their direct subblocks (at depth $d+1$) are disjoint, so the claim follows from the induction hypothesis.
\end{proof}

>>>>

\begin{lemma} \label{lem::subblocksDistinct}
If $B$ and $B'$ are two \rev{different} blocks in nodes at the same depth in the ordering tree, their sets of subblocks are disjoint.
\end{lemma}
\begin{proof}
We prove the lemma by reverse induction on the depth.
If $B$ and $B'$ are in leaves, they have no subblocks, so the claim holds.
Assume the claim holds for nodes at depth $d+1$ and let $B$ and $B'$ be two blocks in nodes at depth $d$.
Consider the direct subblocks of $B$ and $B'$ defined by~(\ref{defsubblock}).
If $B$ and $B'$ are in different nodes at depth $d$, then their direct subblocks are disjoint.
\rev{Suppose $B$ and $B'$ are in the same node $v$.
By Lemma \ref{lem::headProgress}, the $\fld{end\sub{left}}$ values of blocks of $v$ are non-decreasing,
so the intervals of direct subblocks of different blocks in $v$ in $v$'s left child defined by (\ref{defsubblock}) are non-overlapping.
Similarly, the intervals of direct subblocks in $v$'s right child are non-overlapping.
In all cases, the direct subblocks of $B$ and $B'$} (at depth $d+1$) are disjoint, so the claim follows from the induction hypothesis.
\end{proof}
